\chapter{Acknowledgment}

This work would not have been possible without the help of many great people, who I want to thank for their 
participation and support. First of all, my gratitude goes to my supervisors Professor Dr. Stefan Radicke 
and Dr. Andreas Stiegler for supporting my work from the day I approached them. Thank you for providing 
valuable input and discussing topic ideas while always giving me the freedom to pursue my interests. In 
this context I also want to thank everyone who helped me achieve my goals during my studies and all the 
people at Hochschule der Medien for making this possible. \\

\noindent
I want to specifically thank my external supervisor Wilhelm Lorin Atzberger, who I met at Nordic Game Jam 
in April 2024. I never expected finding success in my search for external support just a few weeks before 
starting to work on the thesis. I want to thank you for being spontaneous, for listening and for your 
extraordinary support. Thank you for spending so many days of your time to discuss topics and help me 
reflcet on my implementation. This work would not have been the same otherwise. \\

\noindent
Finally, I want to thank my dear friends who supported my endeavor in endless patience by proof reading and 
providing invaluable feedback in any aspect of the work. I would like to thank Rike Ziegler for always 
supporting me and finding time to read through my drafts multiple times, even though you had your own 
pile of work to do in parallel. Thank you for listening to me, no matter if I was stuck or if I succeeded. 
I want to thank Nikolai Thees for proof reading and always providing great feedback, especially concerning 
the intricacies of the English language. Thank you for your ability to take my spontaneous calls and for 
providing critical and valuable feedback. And I want to thank Malte, not only for being responsable for 
meeting Lorin in the first place, but also for the critical feedback and the invaluable input you provided 
for this work. Thank you for enriching my work with your knowledge and expertise. \\
