\chapter{Motivation} \label{cpt-motivation}


Ever since the early days of computer graphics, both hard- and software have rapidly evolved
alongside the creative and challenging use cases provided by developers, scientists and others.
Many technical increments relied on innovation, especially the advent of new capable hardware.
One example is the \ac{GPU} itself, which is now considered the heart of modern graphics 
processing and is widely used in fields like science, artificial intelligence, games, and pretty 
much all graphics related applications. Before 1995, there already had been a lot of iterations 
on specialized graphics hardware, often focused on video formatting or color output operations
\cite{Singer2023}. [@TODO: Last sentence is redundant]\\

\noindent
The first graphics cards were specialized chips either used for video encoding and decoding
or expensive hardware targeted for large companies, which arised during the early stages 
of "3D consumer graphics" \cite{Singer2023}. Dedicated and affordable \ac{GPU}s for consumer 
\ac{PC} had their large breakthrough during the 1990s. In 1999, NVIDIA introduced their 
first consumer graphics chip, the \emph{NVIDIA GeForce 256}, which was used for efficient 
transformations and lighting \cite{Fenno2024}. Back then, this new hardware created a multitude 
of new possibilities. 
Although it was mostly used for graphics processing - hence the name - nowadays, \ac{GPU}s
are used in a much more general way than in [@TODO: insert year of GPU advent]. Both the 
technical complexity and the fields of application have tremendously increased. Today, a 
\ac{GPU} is often also referred to as a \ac{GPGPU} instead because of its evolution towards 
a multi-purpose tool. \\

\noindent
Nevertheless, \ac{GPU} architecture is still under a strong influence of the entertainment 
industry, first and foremost the games industry. Some of the latest changes to \ac{GPU} 
hardware and \ac{API} design correlates to the ongoing demand for higher output resolutions, 
higher geometric density or additionaly technology for \emph{Deep Learning} algorithms. 
To provide more context for how we made use of some of the more modern features of \ac{GPU}s 
we will give a brief overview over the trends in computer graphics over the past decades, and 
especially the last years.


\section{The Rise Of The GPU}

[@TODO: Focus on the GeForce 256 as the first acclaimed customer graphics card]
Before the integration of specialized hardware, games like \emph{Wolfenstein 3D} or the 
original \emph{Doom} made use of \ac{CPU} Rasterization, with all rendering operations 
executed sequentially by the \ac{CPU} [@TODO: check games and facts].
[@TODO: Check if rendering ops were ALL on CPU] Therefore, a higher poly count usually 
led to lower framerates.

\subsection{Rendering Pipeline}

With the first \ac{GPU} available to larger audiences, games were able to add more and more 
detail into their 3D environment. The heavily parallelized transformations allowed for a lot 
more triangles in games, and the hardware accelerated lighting computations which made for 
even more realistic lighting throughout the rendered scene. Having a dedicated hardware for 
highly demanding tasks started a new era of computer graphics. One significant advantage: 
The \ac{GPU} could operate on a large amount of data while not stalling the \ac{CPU} \cite{Fenno2024}.\\

\noindent 
This new approach of "outsourcing" specific operations quickly became the standard procedure and 
lead to a standardized rendering pipeline. 



\subsection{Deferred Rendering}



\subsection{GPU Driven Rendering}


\section{Hier kommt meine Motivation und Überleitung zur Forschungsfrage}

The latest innovations in computer graphics have been developing in different directions,
one of them being \ac{GPU} Driven Rendering. Minimizing dependencies between \ac{CPU} and \ac{GPU} 
has been an ongoing effort for the last decade. Many of the novel approaches which have been developed 
over the years are visible in the latest technology. But there are still plenty of use cases 
which are not yet living up to their full potential, considering modern hard- and software solutions.\\

\noindent
One of those use cases is Voxel Rendering. Although there are ongoing efforts to improve voxel 
rendering performance, we believe that new innovations can be applied to this field of computer 
graphics. To highlight the possible improvements, one can consider different applications of 
voxel rendering. 






\section{The Idea}

[@TODO: Mention initial idea]

\noindent
In the next chapter we will focus on the first use case, the Rasterized Voxel Rendering. We belive that this 
technique can benefit the most from advances in rendering technology.
With most of the latest improvements of graphics hard- and software, the capabilities to process mesh data 
and compute a rasterized image got better and better. Even the graphics hardware vendor's efforts to enable 
real-time ray traced lighting, shadows, reflections and ambient occlusion is applied within the rasterized 
pipeline and therefore relies on improvements to said pipeline over the last years. \\
We will first lay out a technical foundation in chapter \ref{cpt-technical-background} before discussing 
related work on which our approach relies on in chapter \ref{cpt-related-work}.



- Spectrum between CPU computation and GPU computation. 
CPU <-----------> GPU

- Mesh-Shading pipeline as new \st{standard} pipeline

- Poses question, can mesh shading further optimize voxel rendering?

- Voxel Rendering

- Better geometry creation (in contrast to Geometry Shader)
- Optimized occlusion culling using per-meshlet OC
