\chapter{Case Study} \label{cpt-case-study}

% @TODO: Meshlet Creation -> Voxel

- Meshlet creation for voxel meshes is trivial since meshlets don't have to be per-computed
- Meshlet culling is also straight forward since Voxels are inherently AABBs
- 

\section{Setup}

% @TODO: Add reference to Diligent Engine's github etc.
For the case study, a starting point is used which provides a reliable and comparable environment.
The framework used is the \emph{Diligen Engine} which already implements state of the art, multi-
platform and multi-API rendering. \\

\noindent



// --------------------------------------------------------------------------------- //
Motivation:
Use Meshlet Culling in voxel rendering to optimize GPU Driven Voxel Rendering


Implementation Considerations:

- Meshlet Culling at first, then problem, because meshlet culling algorithm depends on different algos
    - Meshlet Backface Culling -> Not possible because Meshlets are voxels in this case
    - Meshlet Occlusion Culling -> Using 2 Pass Depth Occlusion Culling sounds good!
    - 2PDOC relies on drawing best occluders which are usually hand picked by artists (buildings)
        -> We can dynamically pick best occluders by selecting full octree nodes as best occluders!

- Optimize octree by using High Res SVOs (https://www.cse.chalmers.se/~uffe/HighResolutionSparseVoxelDAGs.pdf).
    - As shown in the paper, here we also only need to encode if a node is filled or not. This way we can 
    optimize memory and proccessing speed! (Not done here because of limitation in loading and voxelizing models.)


- Align octree bounds with outer most voxel bounds, so no misalignment happens!
    
- Limitations: 
    - Using Mesh Shading, we occlude meshlets against other meshlets, instead of instances against other
    instances. This might be a limiting factor!

- Worse performance on slopes? -> Check and measure!

- Larger nodes which are not being split can be "full" without being full!

// --------------------------------------------------------------------------------- //
